Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-09-23 document/ms.tex
update to abstract from Garrett

\begin{abstract}
Ages and masses of young stars can be estimated by comparing observed luminosities and effective temperatures to pre-main sequence stellar evolutionary model tracks.  Nonstandard physics like magnetic fields and starspots confound interpretations of pre-main sequence HR diagrams and their age and mass estimates.  
We characterize the fundamental properties of the spotted weak-lined \emph{T-Tauri} star LkCa 4 by searching for signatures of the emergent radiation originating from the starspot or starspot groups.  The spectral and photometric properties of this radiation can be used to constrain both the starspot filling factor and the starspot temperature.
A generalization of line depth ratio analysis is developed and applied to a high-resolution near-IR IGRINS spectrum.  The IGRINS spectrum is forward modeled with pre-computed \PHOENIX\ synthetic spectra using the open source spectral inference framework \texttt{Starfish} extended to include a two component mixture model comprised of ambient photosphere and starspot emission.  
Spectral features attributable to both ambient photosphere $\teffa\sim4100$ K and cool starspots $\teffb\sim2700-3000$ K are detected in the IGRINS spectrum.  In some spectral regions, the starspot emission acts merely as a relatively featureless ``veiling'' continuum owing to the large rotational broadening ($\vsini\sim28$ km~s$^{-1}$) and heavy line-blanketing in cool star spectra.  The filling factor of of starspots at minimum light $\Delta V = 0.5$must be $>37\%$, and is likely much larger based on geometry and temperature considerations.  The IGRINS spectra yield consistently large ($>50\%$) fill factors across all IGRINS spectral orders, with a best fit filling factor of $f\sim80\%$ starspots at the epoch of spectrum acquisition.  The interpretation of the $\sim20\%$ ambient photosphere as a persistent ``hot spot'' is also possible.  The spectral energy distribution, color variability, TiO variability, rotational modulation, and low-resolution optical/IR spectroscopy show broad support for the starspot parameter estimates derived separately from IGRINS, strengthening the case for a large albeit secularly changing starspot filling factor and temperature contrast through time.
The revised effective temperature and luminosity make LkCa 4 appear much lower mass and much younger compared to previous estimates based on optical spectroscopy and not taking into account the large covering fraction of the cool spots.  Emission from longitudinally symmetric starspots has likely caused the misplacement of spotted young stars on the observational HR diagram, leading to incorrect estimates for ages and masses and contributing, in part, to the apparent age spread observed towards young clusters.
\end{abstract}

>>>>

\begin{abstract}
%Ages and masses of young stars can be estimated by comparing observed luminosities and effective temperatures to pre-main sequence stellar evolutionary model tracks.  Nonstandard physics like magnetic fields and starspots confound interpretations of pre-main sequence HR diagrams and their age and mass estimates.  
%We characterize the fundamental properties of the spotted weak-lined \emph{T-Tauri} star LkCa 4 by searching for signatures of the emergent radiation originating from the starspot or starspot groups.  The spectral and photometric properties of this radiation can be used to constrain both the starspot filling factor and the starspot temperature.
%A generalization of line depth ratio analysis is developed and applied to a high-resolution near-IR IGRINS spectrum.  The IGRINS spectrum is forward modeled with pre-computed \PHOENIX\ synthetic spectra using the open source spectral inference framework \texttt{Starfish} extended to include a two component mixture model comprised of ambient photosphere and starspot emission.  
%Spectral features attributable to both ambient photosphere $\teffa\sim4100$ K and cool starspots $\teffb\sim2700-3000$ K are detected in the IGRINS spectrum.  In some spectral regions, the starspot emission acts merely as a relatively featureless ``veiling'' continuum owing to the large rotational broadening ($\vsini\sim28$ km~s$^{-1}$) and heavy line-blanketing in cool star spectra.  The filling factor of of starspots at minimum light $\Delta V = 0.5$must be $>37\%$, and is likely much larger based on geometry and temperature considerations.  The IGRINS spectra yield consistently large ($>50\%$) fill factors across all IGRINS spectral orders, with a best fit filling factor of $f\sim80$\% starspots at the epoch of spectrum acquisition.  The interpretation of the $\sim20$\% ambient photosphere as a persistent ``hot spot'' is also possible.  The spectral energy distribution, color variability, TiO variability, rotational modulation, and low-resolution optical/IR spectroscopy show broad support for the starspot parameter estimates derived separately from IGRINS, strengthening the case for a large albeit secularly changing starspot filling factor and temperature contrast through time.
%The revised effective temperature and luminosity make LkCa 4 appear much lower mass and much younger compared to previous estimates based on optical spectroscopy and not taking into account the large covering fraction of the cool spots.  Emission from longitudinally symmetric starspots has likely caused the misplacement of spotted young stars on the observational HR diagram, leading to incorrect estimates for ages and masses and contributing, in part, to the apparent age spread observed towards young clusters.

Ages and masses of young stars can be estimated by comparing their observed luminosities and effective temperatures to pre-main sequence stellar evolutionary model tracks, but nonstandard physics like magnetic fields and starspots complicate this process. In order to understand their influence, we study in detail the heavily-spotted weak-lined \emph{T-Tauri} star LkCa 4 by searching for spectral signatures of radiation originating from the starspot or starspot groups. To do this, we introduce new methodology for constraining both the starspot filling factor and the typical spot temperature, which involves fitting two-temperature stellar atmosphere models constructed from \PHOENIX\ synthetic spectra to a high-resolution near-IR IGRINS spectrum. We detect spectral features attributable to both an ambient photosphere $\teffa\sim4100$ K and cool starspots $\teffb\sim2700-3000$ K, and derive consistently large ($>50\%$) fill factors across all IGRINS spectral orders, with a best fit filling factor of $f\sim80$\% starspots at the epoch of spectrum acquisition. These results are bolstered by the spectral energy distribution, color variability, TiO variability, rotational modulation, and low-resolution optical/IR spectroscopy, which collectively provide broad support for the picture of an extremely heavily spotted star with secularly changing starspot filling factors and temperature contrasts through time. A persistent hotspot covering $\sim20$\% of the surface is a plausible alternative explanation. The revised effective temperature and luminosity make LkCa 4 appear much younger and lower mass when compared to previous measurements, a bias likely afflicting most young, heavily spotted stars, and which may contribute to observed age spreads in open clusters. We also discuss the detailed influence of starspots on the IGRINS spectrum, which in some regions acts as a relatively featureless ``veiling'' continuum owing to the large rotational broadening ($\vsini\sim28$ km~s$^{-1}$) and heavy line-blanketing in cool star spectra.

\end{abstract}
